---
title: "GAPMINDER | LIFE EXPECTANCIES 2007"
format: dashboard
theme: lux
---

```{r}
# Load packages
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, 
               gapminder,
               plotly, 
               reactable, 
               bslib, 
               bsicons, 
               shiny,
               rnaturalearth,
               countrycode,
               htmltools,
               reactable)
```

```{r}
gap_07 <- gapminder %>% 
  filter(year == 2007)

highest_life_exp_country <- 
  gap_07 %>% 
  arrange(-lifeExp) %>% 
  head (1) %>% 
  pull(country)

highest_life_exp_value <- 
  gap_07 %>% 
  arrange(-lifeExp) %>% 
  head (1) %>% 
  pull(lifeExp) %>% 
  round(1)

lowest_life_exp_country <- 
  gap_07 %>% 
  arrange(lifeExp) %>% 
  head (1) %>% 
  pull(country)

lowest_life_exp_value <- 
  gap_07 %>% 
  arrange(lifeExp) %>% 
  head (1) %>% 
  pull(lifeExp) %>% 
  round(1)

#for the weighted average life-expectancy

average_life_exp_value <- 
  gap_07 %>% 
  summarise(weighted_mean = sum(lifeExp * pop/ sum(pop))) %>% 
  round(1)
  

```

```{r}
country_shapes <- rnaturalearth::ne_countries()

# ggplot(country_shapes) +
#  geom_sf()

# country_shapes

#to add a column with the ISO codes of each country

gap_07 <- gap_07 %>% 
  mutate(
    country_code = countrycode(country, "country.name", 
"iso3c"))

# to check if there are missing data in the gap_07 dataset column on country code and the country-level shapefile 

# setdiff(gap_07$country_code, country_shapes$adm0_a3)

# to join the datasets, getting the data from gap_07 to ensure it will be reflected in the shapefile 

gap_map <- left_join(country_shapes, gap_07,
          by = c("adm0_a3" = "country_code")) %>% 
  mutate(tooltip_label = paste(country, round(lifeExp, 1), sep = ": ")) %>% 
  ggplot() +
  geom_sf(aes(fill = lifeExp, text = tooltip_label)) +
  theme_void() +
  theme(legend.position = "none")

gap_map_ggplotly <- ggplotly(gap_map, tooltip = "text")


```
```{r}
top20_countries <- gap_07 %>% 
  arrange(desc(lifeExp)) %>% 
  head(20) %>% 
  mutate(tooltip_label = paste(country, round(lifeExp, 1), sep = ": ")) %>% 
  ggplot(aes(y = reorder(country, lifeExp), 
             x = lifeExp, 
             fill = lifeExp,
             text = tooltip_label)) +
  geom_col() +
  theme(legend.position = "none") +
  geom_text(aes(label = round(lifeExp, 1)),
                nudge_x = -10,
                color = "white"
            ) +
  labs (y = "Country",
        x = "Life Exp")

top20_countries_ggplotly <- ggplotly(top20_countries, tooltip = "text")            
```


# HOME

## Row 1 {height=30%}

```{r}
value_box(
  title = "Highest Life Expectancy",
  value = highest_life_exp_value,
  showcase = bsicons::bs_icon("heart"),
  theme = value_box_theme(bg = "#518fd6"),
  p(paste0("(", highest_life_exp_country, ")"))
)
```

```{r}
value_box(
  title = "Lowest Life Expectancy",
  value = lowest_life_exp_value,
  showcase = bsicons::bs_icon("thermometer-half"),
  theme = value_box_theme(bg = "#214773"),
  p(paste0("(", lowest_life_exp_country, ")"))
)
```

```{r}
value_box(
  title = "Average Life Expectancy",
  value = average_life_exp_value,
  showcase = bsicons::bs_icon("graph-up"),
  theme = value_box_theme(bg = "#4059A2")
)
```


## Row 2 {height=70%}



### {width=70%}

```{r title = "Map of Countries by Life Expectancy"}
gap_map_ggplotly
```

### {width=30%}

```{r title = "Top 20 Countries by Life Expectancy"}
top20_countries_ggplotly
```

# DOWNLOAD DATA

The data used in this dashboard is shown below and can be downloaded as a CSV.

```{r}

htmltools::browsable(
  tagList(
    reactable(gapminder,
              elementId = "gapminder-table",
              searchable = T,
              filterable = T),
    tags$button("Download as CSV", onclick = "Reactable.downloadDataCSV('gapminder-table')"),

    )
  )

```


# ABOUT

This data comes from the r 'gapminder' package, and is originally sourced from the Gapminder Foundation.

The Gapminder Foundation is a non-profit venture registered in Stockholm, Sweden, that promotes sustainable global development and achievement of the United Nations Millennium Development Goals by increased use and understanding of statistics and other information about social, economic, and environmental development at local, national, and global levels.

Gapminder was founded in 2005 by Ola Rosling, Anna Rosling Rönnlund, and Hans Rosling. The name Gapminder was derived from the "Mind the Gap" warning messages on the London Underground.

An example of one of Hans Rosling's work is shown below:

<iframe width="560" height="315" src="https://www.youtube.com/embed/hVimVzgtD6w?si=EffNpq0lKY7gU5Yv" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>
